---
title: "Week 1 Lab 1"
author: "Insert your name here"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting started

The Behavioral Risk Factor Surveillance System (BRFSS) is an annual
telephone survey of 350,000 people in the United States. As its name
implies, the BRFSS is designed to identify risk factors in the adult
population and report emerging health trends. For example, respondents
are asked about their diet and weekly physical activity, their HIV/AIDS
status, possible tobacco use, and even their level of healthcare
coverage. The BRFSS Web site (<http://www.cdc.gov/brfss>) contains a
complete description of the survey, including the research questions
that motivate the study and many interesting results derived from the
data.

We will focus on a random sample of 20,000 people from the BRFSS survey
conducted in 2000. While there are over 200 variables in this data set,
we will work with a small subset.

We begin by loading the data set of 20,000 observations into the R
workspace. After launching RStudio, enter the following command.

```{r load-data}
source("http://www.openintro.org/stat/data/cdc.R")
```

The data set cdc that shows up in your workspace (the Environment tab on
the top right) is a data matrix, with each row representing a case (one
respondent) and each column representing a variable. R calls this data
format a data frame, which is a term that will be used throughout the
labs.

## Exercise 1

How many cases are there in this data set? How many variables? For each
variable, identify its data type (e.g. categorical, discrete).

```{r exercise01}
source("http://www.openintro.org/stat/data/cdc.R")
glimpse(cdc)
```

## Exercise 2

Create a numerical summary for height and age, and compute the
interquartile range for each. Compute the relative frequency
distribution for gender and genhlth. How many males are in the sample?
What proportion of the sample reports being in excellent health?

```{r exercise02}
# Q2 
## Numerical summary of Health and Age
summary(cdc)
## Interquartile of Health and Age
## - Height: 64 to 70 in.
## - Age: 31 to 57 years.

# Relative frequency distribution
table(cdc$genhlth)/length(cdc$genhlth)
table(cdc$gender)/length(cdc$gender)

# N of males is shown in line 65. N(males) = 9569

# N of excellent health in line 71. P(excellent) = 23.29%
```

## Exercise 3

Create a new object called under23_and_smoke that contains all
observations of respondents under the age of 23 that have smoked 100
cigarettes in their lifetime. Write the command you used to create the
new object as the answer to this exercise.

```{r exercise03}
library(tidyverse)
library(dplyr)
under23_and_smoke = cdc %>%
  filter(age<23) %>%
  filter(smoke100==1)

summary(under23_and_smoke)
```

## Exercise 4

Find the formula for BMI and create a new variable called BMI,
corresponding to each person's BMI.

```{r exercise04}
cdc$bmi=cdc$weight*703/cdc$height^2
```

## Exercise 5

Make a scatterplot of weight versus desired weight. Describe the
relationship between these two variables.

```{r exercise05}
plot(cdc$weight,cdc$wtdesire)
#desired weight increases linearly with weight
```

## Exercise 6

Let’s consider a new variable: the difference between desired weight
(wtdesire) and current weight (weight). Create this new variable by
subtracting the two columns in the data frame and assigning them to a
new column called wdiff.

```{r exercise06}
cdc$wdiff = cdc$wtdesire - cdc$weight
```

## Exercise 7

What type of data is wdiff? If an observation wdiff is 0, what does this
mean about the person’s weight and desired weight. What if wdiff is
positive or negative?

```{r exercise07}

#wdiff is integer data type. If. an observation is 0, the person does not want to gain or lose any weight. if wdiff is negative then the responder wants to lose weight if wdiff is positive, they want to gain weight.
```

## Exercise 8

Describe the distribution of wdiff in terms of its center, shape, and
spread, including any plots you use. What does this tell us about how
people feel about their current weight?

```{r exercise08}
hist(cdc$wdiff, xlim=c(-200,100),main="Distribution of Desired Weight Change",
xlab="Desired Weight Change (lb)", ylab="Number of Respondents",breaks=100)

#wdiff seems to be centered on the lower side of 0. That is, most people wish to reduce their weight slightly (or none). It seems to have a bell curve shape, skewed to a desired reduction in weight (negative value). The fold change is sharper for those wishing to increase their weight.
#Essentially, most people wish to decrease in weight.
```

## Exercise 9

Using numerical summaries and a side-by-side box plot, determine if men
tend to view their weight differently than women.

```{r exercise09}
menwdiff = cdc %>%
  filter(gender=="m")

boxplot(cdc$wdiff~cdc$gender, ylim=c(-200,100))

tapply(cdc$wdiff, cdc$gender, summary)


```
